### 2014-10-23 writeups/data_sources.tex
start some writtings

\documentclass[]{article}

\usepackage{amsmath, amsfonts, amsthm, amssymb}

% bibliography
\usepackage{natbib}
% \bibpunct{(}{)}{;}{a}{}{,}

%opening
\title{}
\author{}

\begin{document}

\maketitle

\begin{abstract}
Discuss my data sources
\end{abstract}

\section{Ethnic fractionalization}

\cite[p. 190]{Gandhi2008}. (1946-2002, but constant across the period). Degree of polarization among a country's ethnic groups. Calculated as $1 - \sum(0.5 - p_i)^2 \times \frac{p_i}{0.25}$, where$p_i$ is the share of the population that belongs to ethnic group $i$, as defined and measured in \cite{Fearon2003} (which also gives a number to a country instead of time series).

\section{Violence}

\section{Other democracies}

Polity IV (1800 - 2012), \verb`polity_binary` is created as \verb`Polity2` $> 0$.

Democracy and Dictatorship (1946 - 2008)

\bibliographystyle{chicago}
\bibliography{library}

\end{document}