### 2014-11-06 data/public/GWF_Autocratic_Regimes_1_2/README.tex
add Makefile, change script to match data folder

``New Data on Autocratic Breakdown and Regime Transitions''

Version 1.2 June 23, 2012

Barbara Geddes, UCLA 
Joseph Wright, Pennsylvania State
Erica Frantz, Bridgewater State

Corresponding author: Joe Wright (josephgwright@gmail.com)

GWF Codebook.pdf		contains definitions coding rules; explanations of difficult cases; and narratives of regime Start and End events

GWF Autocratic Regimes:

	GWF Autocratic Regimes.xls 	contains three tabs: a list of regime cases; tables for the codebook; and TSCS data

	GWFtscs.txt					contains the regime TSCS data in a txt file

	GWFcases.dta				contains the regime cases in a Stata file

	GWFtscs.dta					contains the regime TSCS data in a Stata file

	clean.do						Stata code to transform the list of cases into TSCS data

GWF All Political Regimes:

	GWF All Political Regimes Case List.xls		contains a list of all political regimes that includes autocratic regime types and democracies

	GWF_AllPoliticalRegimes.dta				contains TSCS data on all political regimes (autocratic regime types and democracies)

	GWF_AllPoliticalRegimes.txt				contains TSCS data on all political regimes (autocratic regime types and democracies)

Research for this project is supported by the National Science Foundation (BCS-0904478 and BCS-090463).

Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) 
and do not necessarily reflect the views of the National Science Foundation.