### 2014-11-06 data/public/GWF_Autocratic_Regimes_1_2/UPDATES.tex

Data first made publicly available in October 2012 (Version 1.0)

CORRECTIONS in Feburary 2013 (Version 1.1):

(1) 2/22/13: Corrected Paraguay 1939-1948; it was 1940-1948 now it is 1939-1948; corrected start year, duration spell, and duration time

(2) 2/22/13: Corrected Greece 1948-1967; it was coded democracy from Jan 1 1948 to Jan 1 1967; now correctly coded democracy from Jan 1 1947 to Jan 1 1967. Updated Global case list and Global regimes. (h/t Kerim Can Kavakli)

CORRECTIONS in June 2014 (Version 1.2):

(3) 6/20/14: Corrected Togo 1967-; it was coded democracy from Jan 1 1964 to Jan 1 1967; now correctly coded as autocratic from Jan 13, 1963 coup led by ex-Sgt Eyadema; the May 1963 election of Grunitzkey was not democratic. A New York Times article states: "With only a single slate of candidates yesterday, the result was a foregone conclusion."  ("Grunitzky Is Declared Winner in Togo's Election" New York Times, May 7 1963.)  First (1970, 207 ) writes: ``The army installed a civilian government, presided over by Olympia's political rival, Nicholas Grunitzky. Bodjolle made himself colonel, commander and chief-of-staff, while a former sergeant in the French army, Etienne Eyadema (who, it is widely believed, fired the shots that killed Olympio) became a major in an army expanded to 1,200 men. Two years later Eyadema ousted Bodjolle, to make himself colonel and commander, and subsequently to become Togo's head of state.'' (h/t Tyson Chandler)

(4) 6/20/14: Corrected Iran 1925-1979; previously the clean.do code and thus the GWFtscs.txt & GWFtscs.dta files had this regime incorrectly coded as gwf_party and gwf_monarch; only the latter is correct; all the GWFglobal data was correct (h/t Wonjun Song)

(5) 6/20/14: Corrected El Salvador 1932-1948; previously was military in GWFcases.dta and GWF Autocratic Regimes.xls but should be military-personal; this information was correct in the GWFglobal data (h/t Brett Ashley Leeds)

(6) 6/20/14: Corrected Cambodia 53-70 end date from March 17 to March 18, 1970; also corrected this date for the Cambodia 70-75 start date; GWFglobal data does not contain start dates so was not updated  (h/t John Chen) "Shinaouk Reported Out in a Coup By His Premier". New York Times, March 18, 1970 (page 1): "Prince Norodom Sihanouk... was overthrown today in his absence, the Pnompenh radio announced."

(7) 6/20/14: JW checked that all data has the correct date for end of the Mauritania 60-78 regime: it is July 10, 1978.  There do not appear to be any instances where this date was incorrectly entered as October 7, 1978. (h/t John Chen)