### 2014-11-21 writeups/appendix_modeldiagnostic.tex
finish model description

\section{Appendix A: Bayesian hierarchical model diagnostic}
\label{sec:modeldiagnostic}