### 2014-11-21 writeups/multilevel_result.tex
start putting in results

\section{Result}

$Pr(liec > 0) = 0.8920333$ close to significant.

\begin{figure}[ht]
    \centering
    \includegraphics[width=.75\textwidth]{../fig/liec_posterior}
    \label{fig:liec_posterior}
\end{figure}