### 2014-11-21 writeups/appendix_modeldiagnostic.tex
writing on experiment

\section{Appendix A: Bayesian hierarchical model diagnostic}
\label{sec:modeldiagnostic}

The traceplots show that the Gibbs sampler performs well with good mixing. The posterior density distribution is very close to normal.

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{../fig/mcmc_diagnostic1}
    \label{fig:mcmc_diagnostic1}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\textwidth]{../fig/mcmc_diagnostic2}
    \label{fig:mcmc_diagnostic2}
\end{figure}