### 2014-11-21 writeups/experiment_checking.tex

\subsection{Checking the experimental assumption}

Since the two prompts are randomly sent out to business, we have excellent balance across observable co-variates as shown in \autoref{tab:balance}.

In addition, to guard against concerns that this question may be politically sensitive, leading to systematic bias in non-response, I also check and confirm that firms with the treatment form does not have a different non-response rate ($\textbf{p-value} = 0.12$).

\begin{table}[ht]
\centering
\begin{tabular}{rrrrl}
  \hline
 & Mean treated & Mean control & p-value &   \\ 
  \hline
Limited Liability & 0.54 & 0.57 & 0.09 & * \\ 
  Joint Stock & 0.23 & 0.23 & 0.84 &  \\ 
  Partnership & 0.00 & 0.00 & 0.38 &  \\ 
  Other form of ownership & 0.01 & 0.00 & 0.03 & ** \\ 
  Manufacturing & 0.08 & 0.09 & 0.57 &  \\ 
  Construction & 0.28 & 0.27 & 0.93 &  \\ 
  Service & 0.66 & 0.67 & 0.47 &  \\ 
  Agriculture & 0.10 & 0.08 & 0.01 & ** \\ 
  Mining & 0.02 & 0.03 & 0.09 & * \\ 
  Revenue (0.5-1 billion VND) & 0.22 & 0.22 & 0.98 &  \\ 
  Revenue (1-5 billion VND) & 0.43 & 0.43 & 0.71 &  \\ 
  Revenue (5-10 billion VND) & 0.07 & 0.08 & 0.23 &  \\ 
  Revenue (10-50 billion VND) & 0.05 & 0.06 & 0.39 &  \\ 
  Revenue (50-200 billion VND) & 0.01 & 0.01 & 0.94 &  \\ 
  Revenue (200-500 billion VND) & 0.00 & 0.00 & 0.81 &  \\ 
  Revenue (above 500 billion VND) & 0.00 & 0.00 & 0.53 &  \\ 
  Size (5-9 people) & 0.36 & 0.34 & 0.12 &  \\ 
  Size (10-49 people) & 0.21 & 0.24 & 0.05 & ** \\ 
  Size (50-199 people) & 0.06 & 0.05 & 0.49 &  \\ 
  Size (200-299 people) & 0.01 & 0.01 & 0.58 &  \\ 
  Size (300-499 people) & 0.00 & 0.00 & 0.20 &  \\ 
  Size (500-1000 people) & 0.00 & 0.00 & 0.81 &  \\ 
  Size (more than 1000 people) & 0.00 & 0.00 & 0.08 & * \\ 
  Main customer (State) & 0.19 & 0.19 & 0.57 &  \\ 
  Main customer (Private) & 0.58 & 0.62 & 0.02 & ** \\ 
  Main customer (Foreign in VN) & 0.05 & 0.04 & 0.33 &  \\ 
  Main customer (Foreign, direct) & 0.03 & 0.03 & 0.25 &  \\ 
  Main customer (Foreign, indirect) & 0.02 & 0.01 & 0.14 &  \\ 
  Equitized local state-owned enterprise & 0.05 & 0.06 & 0.21 &  \\ 
  Equitized central state-owned enterprise & 0.01 & 0.01 & 0.28 &  \\ 
  Firm with some state-owned equities & 0.03 & 0.02 & 0.26 &  \\ 
  Firm that was formerly household business & 0.74 & 0.73 & 0.24 &  \\ 
  Firm with shares listed & 0.01 & 0.00 & 0.39 &  \\ 
  Firm owner graduated from university & 0.62 & 0.63 & 0.91 &  \\ 
  Firm owner with MBA & 0.03 & 0.04 & 0.45 &  \\ 
  Firm owner was a leader of a State agency & 0.04 & 0.03 & 0.37 &  \\ 
  Firm owner was a military officer & 0.05 & 0.04 & 0.26 &  \\ 
  Firm owner was a manager of SOE & 0.13 & 0.12 & 0.22 &  \\ 
  Firm owner was an employee at SOE & 0.15 & 0.15 & 0.87 &  \\ 
   \hline
\end{tabular}
\caption{Covariate balance in survey experiment}
\label{tab:balance}
\end{table}