### 2014-11-21 writeups/experiment_result.tex

\subsection{Result}

Given the experimental design of our survey experiment, estimating the treatment effect is straight forward with a simple logistic regression as reported in \autoref*{tab:survey_result}.

The treatment is highly significant ($p < 0.01$) with a coefficient of 0.264. This means that, when prompted to think about the inclusion of businessmen in the National Assembly, the respondent is 1.3 time more likely to think that participating in economic policy making via the National Assembly is effective.

\begin{table}[!htbp] \centering 
  \caption{Treatment effect in Survey experiment} 
  \label{tab:survey_result} 
\begin{tabular}{@{\extracolsep{5pt}}lc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{1}{c}{\textit{Dependent variable:}} \\ 
\cline{2-2} 
\\[-1.8ex] & Contacting the National Assembly is effective \\ 
\hline \\[-1.8ex] 
 Treatment prompt & 0.264$^{***}$ \\ 
  & (0.073) \\ 
  & \\ 
 Constant & 0.929$^{***}$ \\ 
  & (0.049) \\ 
  & \\ 
\hline \\[-1.8ex] 
Observations & 3,951 \\ 
Log Likelihood & $-$2,253.137 \\ 
Akaike Inf. Crit. & 4,510.274 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{1}{r}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
\end{tabular} 
\end{table}