### 2014-11-22 writeups/appendix_modeldiagnostic.tex
done with all, doing references

\section{Appendix A: Bayesian hierarchical model diagnostic}
\label{sec:modeldiagnostic}

The traceplots show that the Gibbs sampler performs well with good mixing. The posterior density distribution is very close to normal.

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{../fig/mcmc_diagnostic1}
    \label{fig:mcmc_diagnostic1}
\end{figure}

\begin{figure}[H]
    \centering
    \includegraphics[width=0.9\textwidth]{../fig/mcmc_diagnostic2}
    \label{fig:mcmc_diagnostic2}
\end{figure}