### 2014-11-22 writeups/draft.tex

\documentclass[12pt]{article}
% This first part of the file is called the PREAMBLE. It includes
% customizations and command definitions. The preamble is everything
% between \documentclass and \begin{document}.

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments

\usepackage{rotating} % for sideway table
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
\usepackage{cleveref}

% bibliography
\usepackage{natbib}
% \bibpunct{(}{)}{;}{a}{}{,}

\usepackage{array,tabularx}

\newenvironment{conditions*}
  {\par\vspace{\abovedisplayskip}\noindent
   \tabularx{\columnwidth}{>{$}l<{$} @{${}={}$} >{\raggedright\arraybackslash}X}}
  {\endtabularx\par\vspace{\belowdisplayskip}}

\usepackage{float}
\restylefloat{table}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

\title{The effect of authoritarian legislature on regime-opposition interaction}
\author{Anh Le}

\begin{document}
\maketitle

\input{introduction}
\input{argument_overview}
\input{theory}

\section{Cross-country analysis of political violence using event data}
\label{sec:empirics_largeN}

\input{multilevel_datadescription}
\input{multilevel_empiricalstrategy}
\input{multilevel_model}
\input{multilevel_result}

\section{Survey Experiment}
\label{sec:empirics_survey}

\input{experiment_description}
\input{experiment_checking}
\input{experiment_result}

\input{conclusion}

\pagebreak
\appendix
\input{appendix_modeldiagnostic}

\pagebreak
\bibliographystyle{chicago}
\bibliography{library}

\end{document}