### 2014-11-22 writeups/introduction.tex

\section{Introduction}
\label{sec:introduction}

Once disregarded as mere ``window-dressing,'' authoritarian institutions have garnered the attention of scholars as an important tool for power sharing. These institutions are found to have various substantive effects, including regime survival, increased investment, and better economic performance \citep{Gandhi2008, Wright2008, Boix2013, Jensen2012}.

While our understanding of authoritarian institution has branched to numerous substantive topics, its causal root deserves more careful studies. Given its common definition as the ``rule of the game,'' institutions are fundamentally a set of constraints on human interactions, and thus its effect must be most evident in the interaction between strategic actors itself \citep{North1990}. And yet most studies have focused on the secondary socioeconomic effects of authoritarian legislature instead of investigating the dynamics of regime-opposition interaction, which is causally prior \citep{Gandhi2007, Svolik2012}. The goal of this article is, therefore, to empirically examine this interaction, arguing that authoritarian legislature should reduce poliltical violence by providing information, lengthening the bargaining game, and enabling side payments.

Beyond its theoretical contribution, a study of violence in partially liberalized dictatorship has immense real-world importance. Indeed, violence during periods of political liberalization is the greatest concern of many would-be democrats living under authoritarian regimes. For example, Chinese citizens show overwhelming enthusiasm when asked about democratic features such as popular accountability, separation of power, and political liberalism, yet remain consistently cautious of instability caused by transition \citep[309]{Chu2008}. Therefore, understanding the pattern of violence during liberalization is crucial for the prospect of democracy.

The article will proceed as follows. \Cref{sec:argument} presents an overview of the theoretical argument and empirical analyses. \Cref{sec:theory} posits two theoretical explanations for the effect of legislature on violence. \Cref{sec:empirics_largeN} uses a Bayesian hierarchical model to analyze political violence in authoritarian regimes post-Cold War, showing that the legislature has a modest positive effect. \Cref{sec:empirics_survey} uses a survey experiment to demonstrate that co-optation strategy does prompt Vietnamese business to interact with the regime, a pre-requisite for any material outcome such as violence to change. \Cref{sec:conclusion} concludes.