### 2014-11-22 writeups/multilevel_model.tex

\subsection{The model}

Given that my dependent variable of interest (\textit{violence}) is at the event level, while my ``treatment'' (\textit{legislature}) is at the country year level, a Bayesian hierarchical model as the most appropriate choice.\footnote{The choice of Bayesian hierarchical model is largely practical: with 15,037 events, popular R packages to fit frequentist hierarchical model do not converge due to a high chance of numerical issues.}  Each event belongs to a country-year, and each country-year belongs to a country. The intercept at each level varies according to the group the unit belongs to. This model choice is superior to the alternative option, which is to aggregate the event-level data up to the country-year level and run a fixed effect model. Doing so will 1) ignore the variation at the event level, resulting in standard errors that are too small and thus false positives; 2) leave the aggregation mechanisms up to the researcher with no clear theoretical reasoning.\footnote{Should we measure the level of violence within a country year by averaging across events? Why not the median? Or count the number of positive / negative events? The choice is not likely to be theoretically driven and open to abuse.}

All parameters have conjugate un-informative prior so that the result is driven entirely by the data instead of the prior.\footnote{More specifically, the prior for coefficient estimates is $Normal(\text{mean}=0, \text{precision}=.0001)$. The prior for the variance parameters is $Inverse-Gamma(1, 1)$} The model is thus:

\begin{alignat*}{2}
&{event}_i &&\sim N(\alpha^{countryyear}_{j[i]} + A \cdot \text{sector}_i, \sigma^2_{event}) \\
&\alpha_j^{countryyear} &&\sim N(\beta_{k[j]}^{country} + \gamma \cdot \text{legis}_j + \Gamma X_j, \sigma_{\alpha}^2) \\
&\beta_k^{country} &&\sim N(\delta_0 + \delta_{ethnic} \cdot \text{ethnic}_k + D \cdot \text{region}, \sigma_\beta^2) 
\end{alignat*}

At the event level:

\begin{conditions*}
\alpha^{countryyear}_{j[i]}    &  intercept of $countryyear_j$ to which $event_i$ belongs\\
\text{sector}_i     &  matrix of dummies for the dissident sector of $event_i$  \\

\end{conditions*}

At the country year level:
\begin{conditions*}
\beta_{k[j]}^{country} &  intercept of $country_k$ to which $countryyear_j$ belongs. In traditional panel data model, this is the country fixed effect. \\
\text{legis}_j & dummy variable for legislature. This is our parameter of interest. \\
X_j & matrix of other control variables, including log(GDP), log(GDP per capita), military expenditure (as \% of GDP), oil and mineral export (as \% of GDP) from \citet{WorldBank2012}; regime duration (years), and authoritarian typology (military, personal, and party as coded by \citet{Geddes2014})
\end{conditions*}

At the country level:
\begin{conditions*}
\text{ethnic} & level of ethnic fractionalization as defined by \citet{Fearon2003} \\
\text{region} & matrix of regional dummies
\end{conditions*}

I fit this model with a Gibbs sampler written in JAGS, running three chains, each has 11,000 iterations with 1,000 burn-ins. The mixing in the traceplot is excellent (i.e. having little autocorrelation) and does not vary across three chains, meaning that our sampling process has accurately approximated the true posterior distribution of the model's parameters \citep{Gelman2006}. (See \autopageref{sec:modeldiagnostic} for more model diagnostics).